%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Welcome to Overleaf --- just edit your LaTeX on the left,
% and we'll compile it for you on the right. If you give
% someone the link to this page, they can edit at the same
% time. See the help menu above for more info. Enjoy!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title{Math 453 HW 1}
\documentclass[addpoints]{exam}

\usepackage{amsmath,enumitem,wrapfig}
\usepackage{tikz}

\newcommand{\StudentName}{Radiation Physics}
\newcommand{\AssignmentName}{HW 1}

\pagestyle{headandfoot}
\runningheadrule
\firstpageheadrule
\firstpageheader{Math 453}{\StudentName}{\AssignmentName}
\runningheader{Math 453}{\StudentName}{\AssignmentName}
\firstpagefooter{}{}{}
\runningfooter{}{}{}

\printanswers

\begin{document}


Organize your work and show any work that you want credit for. Use full sentences where possible.

\begin{questions}
    \question \textbf{QUESTION 1 (30 MARKS)}
    \begin{parts}
        \part 
        1.1.
        Explain the following terms
        (a) Compton or recoil electrons, Beta particles and Pair Production electrons [3 marks] 
        (b) Synchrotron radiation and Annihilation quanta [2 marks]
        (c) Remote afterloading technique and Neutron activation [2 marks]
        \begin{solution}
        A Compton/Recoil electron refers to an electron that is ejected from an atom during the Compton scattering process. Compton scattering is a phenomenon in quantum physics where a photon (usually an X-ray or gamma-ray photon) collides with an electron, and as a result of the collision, the photon loses energy, and the electron is ejected from its orbit around an atomic nucleus.
        
A beta particle, also known as a beta ray or beta radiation, is a high-energy, high-speed electron or positron emitted by the radioactive decay of an atomic nucleus during the process of beta decay.A beta electron refers to an electron that is emitted during a type of radioactive decay known as beta decay. Beta decay involves the transformation of a neutron into a proton or a proton into a neutron within an atomic nucleus, accompanied by the emission of a beta particle.
        Synchrotron radiation refers to electromagnetic radiation emitted by charged
particles following a curved trajectory in free space under the influence of a
magnetic field.
Annihilation quanta refers to the high-energy photons (gamma rays) that are produced when a particle and its antiparticle collide and annihilate each other. During the annihilation process, the total energy and momentum of the initial pair are conserved and distributed among the resulting particles. 
Photons resulting from positron–electron annihilation is the annhiliation quanta. 
The remote afterloading technique is a method used in brachytherapy, a form of radiation therapy, where the radioactive sources are placed within or near the treatment area.
        \end{solution}
        \part 
        1.2. (a) Write a nuclear equation when cesium-137 decays into barium-137. [3 marks] 
        (b) Calculate the decay energy Q- for the B- decay of cesium-137 into barium-137. [5mks] 
        (c) When Fluorodeoxyglucose (FDG) is labelled with radionuclide fluorine-18 it can be injected intravenously into a patient for use in positron emission tomography (PET) functional imaging. What are some of the uses of the FDG PET? [4 marks]
        \begin{solution}
            
        \end{solution}
        \part 1.4. Briefly explain directly ionizing radiation and indirectly ionizing radiation. Give examples for each. [6 marks]
       \begin{solution}
        
       \end{solution}
       \part 1.3. Write down the most important characteristics of radionuclides used in external beam radiotherapy. [5 marks]

    \end{parts}
    
\end{questions}

\begin{questions}
    \question \textbf{Question 2 (20 marks)}
    \begin{parts}
        \part 
        2.1. Which physical quantities are conserved during a nuclear transformation? [5 marks]
        \begin{solution}
            
        \end{solution}
        \part 2.2. The figure below shows an energy level diagram for
        (a) Explain the decay scheme shown above as it ends at the ground state. [5 marks] Calculate the decay energy (Q) for the x decay of 228Ra (radium). [5 marks] Calculate the kinetic energy Ex1 of the x- particle.

        (b) Calculate the decay energy \((Q_B)\) and \((Q_EC)\) for the decay of iridium-192 [8 marks]
        (c) Explain Briefly shy iridium-192 is more widely used in brachytherapy rather than the traditional radium or caesium. What is the half-life of iridium-192? [3 marks]
        
    \end{parts}
        
\end{questions}

\begin{questions}
    \question \textbf{Question 3 (20 marks)}
    \part 3.1. Electron Capture and B+ are competing processes. Write down conditions that are required for one process to proceed over the other process. [4 marks]
    \newline Electron capture is an alternative to B+ decay. Comparison of Eqs. 3.61 and 
3.66 shows that nuclei that cannot experience B+ decay can undergo EC, since 
a smaller mass difference is required for the latter process. Of course, EC is 
always possible if B+ decay is decays both by B+ and EC.
    \begin{solution}
        
    \end{solution}
    \part 3.2 Discuss any two ways or mechanisms by which photons can interact with matter that play a very important role in therapeutic as well as in diagnostic medical physics. [8 marks]
    \begin{solution}
        
    \end{solution}
    \part 3.3. An X-ray photon beam of photon energy 15 eV is allowed to fall on a metal surface. If the threshold frequency for the metal is \(1.2 × 1015 Hz\) , then find the maximum kinetic energy gained by a photoelectron, the maximum speed of a photoelectron, and the stopping potential corresponding to the maximum kinetic energy.
    [8 marks]
    \begin{solution}
        
    \end{solution}        
\end{questions}


\begin{questions}
    \question \textbf{QUESTION 4 (20 MARKS)}
    \part 
    4.1. Explain the following terms. (a) the linear attenuation coefficient and Mass attenuation coefficient [2 marks] (b) First half-value layer HVL,, the second half-value layer HVL2 [2 marks]
    \begin{solution}
        
    \end{solution}
    \part 4.2. Derive the functional relationship shown below between the activity of a radioactive substance and time. \begin{equation}
        A = A_oe-2z
    \end{equation}
    \part 4.3. Draw a typical curve for the radioactive substance as given by the expression in Question 4.2. Briefly explain how you will determine the 2. [6 marks]
    \begin{solution}
        
    \end{solution}
    \part 4.4. If a 0.2-cm thickness of material transmits 25% of a monoenergetic beam of photons, calculate the half-value layer (HVL) of the beam for that material.
    [2 marks]
    \begin{solution}
        
    \end{solution}
\end{questions}

\begin{questions}
    
    \question \textbf{QUESTION 5 (20 MARKS)}
    \part 5.1. Describe the term stopping power of a material.[2]
    \part 5.2. Explain the following quantities that are used for the purpose of quantifying radiation. (a) Activity A, (b) Kerma (K), (c) Equivalent Dose (H)[6]
    \begin{solution}
        
    \end{solution}
    \part 5.3 Write down factors that are considered to choose a radiation beam and dose prescription in treatment of disease with radiation[4]
    \begin{solution}
        
    \end{solution}
    \part5.4 Draw and explain the Bragg curve. What is the importance of the enhanced ionization in the Bragg peak?[8]
    \begin{solution}
        
    \end{solution}
\end{questions}

\begin{questions}
    \question \textbf{QUESTION 1 (30 MARKS)}
    \begin{parts}
        \part 
        1.1.
        Explain the following terms
        (a) Compton or recoil electrons, Beta particles and Pair Production electrons [3 marks] 
        (b) Synchrotron radiation and Annihilation quanta [2 marks]
        (c) Remote afterloading technique and Neutron activation [2 marks]
        \begin{solution}
        
        \end{solution}
        \part 
        1.2. (a) Write a nuclear equation when cesium-137 decays into barium-137. [3 marks] 
        (b) Calculate the decay energy Q- for the B- decay of cesium-137 into barium-137. [5mks] 
        (c) When Fluorodeoxyglucose (FDG) is labelled with radionuclide fluorine-18 it can be injected intravenously into a patient for use in positron emission tomography (PET) functional imaging. What are some of the uses of the FDG PET? [4 marks]
        \begin{solution}
            
        \end{solution}
        \part 1.3. Which physical quantities are conserved during a nuclear transformation[5 mks]
        \begin{solution}
            
        \end{solution}
        \part 1.4. Briefly explain directly ionizing radiation and indirectly ionizing radiation. Give examples for each. [6 marks]
       \begin{solution}
        
       \end{solution}
       
    \end{parts}
    
\end{questions}

\begin{questions}
    \question \textbf{QUESTION 2 (20 MARKS)}
    \begin{parts}
        \part 2.1. Write down the most important characteristics of radionuclides used in external beam radiotherapy. [4 marks] 
        \begin{solution}
            
        \end{solution}
        \part 2.2.The figure below shows an energy level diagram for 228Ra (radium) decaying through a-decay into 22Ra (radon).
        \begin{solution}
            
        \end{solution}
        \part (a)Explain the decay scheme shown above as it ends at the ground state.[5 marks]
        \begin{solution}
            
        \end{solution}
        \part (b)Calculate the decay energy (Q.) for the decay of 228Ra (radium).[5 marks]
        \begin{solution}
            
        \end{solution}
        \part (c)Calculate the kinetic energy E.1 of the- particle.[3 marks]
        \begin{solution}
            
        \end{solution}
        \part (d) Explain why modern brachytherapy is now carried out with other radionuclides (e.g., iridium-192, cesium-137, iodine-125, etc.) instead of the radium-226 which was very popular in the past century?[3 marks]
        \begin{solution}
            
        \end{solution}
    \end{parts}
\end{questions}


\begin{questions}
    \question \textbf{Question 3 (20 marks)}
    \begin{parts}
        \part 3.1. Briefly describe the areas where Ionizing radiation is used. [8 marks]
        \begin{solution}
        
        \end{solution}
        \part 3.2 Electron Capture and B are competing processes. Write down conditions that are required for one process to proceed over the other process. [4 marks]

        \begin{solution}
            
        \end{solution}
        \part 3.3. Discuss any two ways or mechanisms by which photons can interact with matter that play a very important role in therapeutic as well as in diagnostic medical physics. [8 marks]
        
    \end{parts}
    
\end{questions}

\begin{questions}
\question    \textbf{Question 4 (20 marks)}
\part 4.1. Explain the terms linear attenuation coefficient and mass attenuation coefficient [2 marks]
\begin{solution}
    
\end{solution}
\part 4.2. Derive the functional relationship shown below between the activity of a radioactive substance and time. Explain the terms.
\(A = A_oe-yt\) [7 marks]
\begin{solution}
    
\end{solution}
\part 4.3. Draw a typical curve for the radioactive substance as given by the expression in 4.2. Briefly explain how you will determine the \(gamma\) [6 marks]
\begin{solution}
    
\end{solution}
\part 4.4. The half-life of strontium-90, (Sr), is 28.8 years. Calculate the (a) decay constant and (b) the initial activity of 1.00g of the material. [3 marks]
\begin{solution}
    
\end{solution}
\end{questions}

\begin{questions}
    
    \question \textbf{QUESTION 5 (20 MARKS)}
    \part 5.1. Describe the term stopping power of a material.[2]
    \part 5.2. Explain the following quantities that are used for the purpose of quantifying radiation. (a) Activity A, (b) Kerma (K), (c) Equivalent Dose (H)[6]
    \begin{solution}
        
    \end{solution}
    \part 5.3 Write down factors that are considered to choose a radiation beam and dose prescription in treatment of disease with radiation[4]
    \begin{solution}
        
    \end{solution}
    \part 5.4 (a) The Bragg curve below shows a plot of the variation of ionization density as a function of distance travelled by an alpha (∞) particle in air. Explain that form of the curve shown below. [6 marks]
    \begin{solution}
        
    \end{solution}
    \part (b) What is the practical application of the enhanced ionization in the Bragg peak? [2 marks]
    \begin{solution}
        
    \end{solution}
\end{questions}

\begin{questions}
    \question \textbf{QUESTION 1 (30 MARKS)}
\part 1.1. Explain the following terms:
(a)Photoelectrons, Thermions and Megavoltage electrons [3 marks]
(b)Characteristic (fluorescence) x-rays and Bremsstrahlung x-rays [2 marks]
(c) Remote afterloading technique and Neutron activation [2 marks]
\begin{solution}
    
\end{solution}
\part 1.2. (a) Write a nuclear equation when fluorine-18 decays into oxygen-18.[3 marks]
\begin{solution}
    
\end{solution}
 \part (b) Calculate the decay energy Q+¿¿ for the B+ decay of fluorine-18 into oxygen-18. (c) When Fluorodeoxyglucose (FDG) is labelled with radionuclide fluorine-18 it can be injected intravenously into a patient for use in positron emission tomography (PET) functional imaging. What are some of the uses of the FDG PET?[5 marks]
 \begin{solution}
    
 \end{solution}
 \part The half-life of strontium-90, (Sr), is 28.8 years. Calculate the (a) decay constant and (b) the initial activity of 1.00g of the material. [5 marks]
 \begin{solution}
    
 \end{solution}
 \part Briefly describe the areas where ionizing radiation is used [6 marks]
 \begin{solution}
    
 \end{solution}
\end{questions}

\begin{questions}
    \question \textbf {QUESTION 2 (20 MARKS)}
    \begin{parts}
        \part 2.1.Which physical quantities are conserved during a nuclear transformation?[5 marks]
        \begin{solution}
            
        \end{solution}
        \part 2.2.
        The figure below shows an energy level diagram for 228Ra (radium) decaying through a-decay into 22Ra (radon).
        (a)
        Explain the decay scheme shown above as it ends at the ground state.
        [5]
        \part (b) Calculate the decay energy (Q.) for the decay of 228Ra (radium).[5]
\part (c) Calculate the kinetic energy E.1 of the- particle.[3]
\part (d) Explain why modern brachytherapy is now carried out with other radionuclides (e.g., iridium-192, cesium-137, iodine-125, etc.) instead of the radium-226 which was very popular in the past century? [2]

    \end{parts}

\end{questions}

\begin{questions}
\question \textbf{QUESTION 3 (20 MARKS)}    
\begin{parts}
    \part 3.1. Write down the possible outcomes when an X-ray beam or gamma radiation passes through an object.[3]
    \begin{solution}
        
    \end{solution}
    \part 3.2. Discuss any two ways or mechanisms by which photons can interact with matter that play a very important role in therapeutic as well as in diagnostic medical physics.[8]
    \begin{solution}
        
    \end{solution}
    \part 3.3. An X-ray photon beam of photon energy 10 eV is allowed to fall on a metal surface. If the threshold frequency for the metal is 1.5 × 1015 Hz, then find the maximum kinetic energy gained by a photoelectron, the maximum speed of a photoelectron, and the stopping potential corresponding to the maximum kinetic energy.[8 marks]
    \begin{solution}
        
    \end{solution}
\end{parts}

\end{questions}

\begin{questions}
    \question \textbf{QUESTION 4 (20 MARKS)}
\begin{parts}
    \part 4.1. Explain the following terms.
    (a) the linear attenuation coefficient and Mass attenuation coefficient[2]
    (b) First half-value layer HVL1, the second half-value layer HVL2 and tenth value layer [3]
    \begin{solution}
        
    \end{solution}
    \part 4.2. Derive the functional relationship shown below between the thickness of an absorber and intensity of a photon beam attenuated by the absorber. Explain the terms.[7]
\(I(x)\)
\begin{solution}
    
\end{solution}
\part 4.4. If a 0.2-cm thickness of material transmits 25% of a monoenergetic beam of photons, calculate the half-value layer (HVL) of the beam for that material.
[4]
\begin{solution}
    
\end{solution}
\end{parts}

\end{questions}

\begin{questions}
    \question \textbf{QUESTION 5 (20 MARKS)}
    \begin{parts}
        \part 5.1. Describe the term stopping power of a material. [2 marks]        
        \part 5.2.Explain the following quantities that are used for the purpose of quantifying radiation. (a) Activity A,
        (b) Kerma (K),
        and (c) Dose (H) [6 marks]
        \part 5.3. Write down the most important characteristics of radionuclides used in external beam radiotherapy [4]
        \part 5.4. (a)  The Bragg curve below shows a plot of the variation of ionization density as a function of distance travelled by an alpha (x) particle in air. Explain that form of the curve shown below. [6 marks]
        (b) What is the practical application of the enhanced ionization in the Bragg peak? [2 marks]

\begin{solution}
    
\end{solution}

    \end{parts}
\end{questions}

\end{document}

